---
title: "dispositions-and-listings"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{dispositions-and-listings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(width=250,tibble.print_min=Inf,tibble.width=Inf)

knitr::opts_chunk$set( 
  collapse = TRUE,
  comment = "#>"
)

#devtools::build_rmd("vignettes/dispositions-and-listings.Rmd") 
```

```{css, echo=FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
```

```{r setup, warning=FALSE}
library(sweets)
library(dmcognigen)
library(dplyr,warn.conflicts = FALSE)

```

::: callout-note
During the data build records are flagged with a deletion flag variable called DELFN and removed. Records that have been removed are stored in the delimp directory. To do the disposition counts the removed records are added back to the data build. For this vignette we are starting at this step where deleted delimp records have already been added back to the built analysis ready dataset.
:::

# 1) Load a dataset

```{r}
data(broken_pk, package = "sweets")
```

## Check the deletion flags.

```{r}

broken_pk %>%
  cnt(DELFN,DELFNC,prop = FALSE)
```

## Check DVID

We want to track and count the removal of **concentration** records, not **dose** records (i.e. DVID=0 & EVID=1). DELFN = 0 are records kept for the analysis ready dataset build. Records with Non zero DELFN were removed, stored in delimp and combined back for the disposition and listings.

```{r}

broken_pk %>%
  cnt(DVID,DVIDC,prop = FALSE)

```

For the deletions that involve dose records, deletion flags are carried from the dose record (DVID=0, EVID=1) onto the affected concentration records (DVID=1) after the dose record of concern. Here we see all the concentration records after the missing dose DTTM have flag 14.

```{r}

broken_pk %>%
    group_by(USUBJID) %>%
    filter(any(DELFN==14)) %>%
    select(USUBJID,DTTM,DVID,TSFD,DELFN)

```

# 2) Subset the data build to concentration i.e. sample records only

```{r}

broken_pk_samples <- 
  broken_pk %>%
  filter(EVID==0)

```

Just concentration records are left

```{r}
broken_pk_samples %>%
  cnt(DVID,DVIDC,prop = FALSE)
```

# 3) Make a disposition table

Here we are splitting by STUDYID

```{r}

disposition_table <- broken_pk_samples %>%
  sweet_disposition(subjid = USUBJID,group_vars = STUDYID) 

disposition_table %>%
  print(n = 15,width = Inf)


```

Someone changed their mind and now we want to keep DELFN 50 (Analyst-Identified Outliers) records in the dataset. Let's set DELFN 50 as a 'count and keep' flag. Notice how the 2 samples that were previously removed are now only 'Affected' and not 'Excluded'

}

```{r}

broken_pk_samples %>%
  sweet_disposition(subjid = USUBJID,
                    group_vars = STUDYID,
                    cnt_n_keeps = 50) %>%
  print(n = 15,width = Inf)

```

## Pool disposition

pooling a disposition with pool_disposition() will ignore any group stratification set in your disposition table and add a Grand Total row.

```{r}

pooled_disposition <- disposition_table %>%
  pool_disposition() 

pooled_disposition %>%
  print(n = 15, width = Inf)


```

## Filtered disposition

For most projects it makes sense to include a "report" version of the disposition table with only DELFN greater than 7

```{r}

report_disposition <- disposition_table %>%
  filter(`Flag #`> 7) 

report_disposition %>%
  print(n = 15, width = Inf)

```

## Write disposition

```{r, echo=TRUE, eval=FALSE}

disposition_tabs <- list("pooled"= pooled_disposition,
                         "report"= report_disposition,
                         "programmer"= disposition_table,
                         )

writexl::write_xlsx(disposition_tabs,"disposition.xlsx")

```

::: {.callout-warning}
It is best practice to refer back to the counts in your code and check the results against your newly generated disposition table
:::

### Troubleshooting

Use `subjects_in_flag` attribute to look for subjects by DELFN

```{r}

attr(disposition_table,"subjects_in_flag") %>%
    filter(DELFN %in% c(50,10)) %>%
    tidyr::unnest(subjects_in_flag)

```

# Deletion Listings

### Setup the report

Initialize the deletion listings report by calling `build_report()`

```{r}

del_report <- build_report(file_name = "pk-deletion-listings")

```

::: {.callout-tip}
Listings are project specific. Always have a discussion about what should be included.
:::

Check the flags and reasons in the disposition with samples excluded \> 0. We don't need empty tables in our listings (e.g. CDISCPILOT02 Missing Concentration Value). These remaining 7 rows indicate there should be 7 tables in our listings.

```{r}
disposition_table %>%
  filter(`Samples Excluded`>0) %>%
  select(STUDYID,`Flag #`,`Reason for Deletion`,`Samples Excluded`)
```

## Add listing tables

Again we take our `broken_pk_samples`dataset consisting of the analysis ready dataset plus the deleted records that have been removed during the build and now added back. Listings are simply a consolidation of one or more tables. These tables need to be customized to show variables that make sense for each deletion reason. Perhaps we would want to show `TRT` or `ACTARM` for **Placebo Group Not Included in the Analysis** or `DTTM` for **Missing Sample Date and/or Time.** Try to use source variables over build variables especially if 

If your dataset has labels, the label will appear as the variable name on the table. To turn this off populate `...` of `add_listing_to_report()` using `use_attributes = 'none'`. See [reporter::create_table()](https://reporter.r-sassy.org/reference/create_table.html) for additional table specifications. If you need a quick way to add, format, or modify an existing label add [reporter::define()](https://reporter.r-sassy.org/reference/define.html) customization to `listing_table()`

```{r}

# call the initialized deletion report 
# build listing tables and add them inside add_listing_to_report()
# Add tables to the report

del_report <- del_report %>%
  add_listing_to_report(
    listing_table(
    broken_pk_samples %>% 
    filter(STUDYID=='CDISCPILOT01', DELFN == 7) %>%
    arrange(STUDYID, USUBJID, DTTM) %>%
    select(STUDYID,USUBJID,TRT),
    title = "Study 01 Placebo Group Not Included in the Analysis")
  ) %>%
  add_listing_to_report(
    listing_table(
    broken_pk_samples %>% 
    filter(STUDYID=='CDISCPILOT02', DELFN == 7) %>%
    arrange(STUDYID, USUBJID, DTTM) %>%
    select(STUDYID,USUBJID,TRT,ACTARM),
    title = "Study 02 Placebo Group Not Included in the Analysis")
  )%>%
  add_listing_to_report(
    listing_table(
    broken_pk_samples %>% 
    filter(STUDYID=='CDISCPILOT01', DELFN == 9) %>%
    arrange(STUDYID, USUBJID, DTTM) %>%
    select(USUBJID,DTTM,VISIT,PCTPT,PCTEST,PCSTRESN),
    title = "Study 01 Missing Concentration Value")
  ) %>%
  add_listing_to_report(
    listing_table(
    broken_pk_samples %>% 
    filter(STUDYID=='CDISCPILOT01', DELFN == 10) %>%
    arrange(STUDYID, USUBJID, DTTM) %>%
    select(USUBJID,DTTM,VISIT,PCTPT,PCTEST,PCSTRESN),
    title = "Study 01 Missing Sample Date and/or Time")
  ) %>%
  add_listing_to_report(
    listing_table(
    broken_pk_samples %>% 
    filter(STUDYID=='CDISCPILOT02', DELFN == 14) %>%
    arrange(STUDYID, USUBJID, DTTM) %>%
    select(USUBJID,DTTM,VISIT,PCTPT,PCTEST,DVIDC,DV),
    title = "Study 02 Missing Sample Date and/or Time")
  )%>%
  add_listing_to_report(
    listing_table(
    broken_pk_samples %>% 
    filter(STUDYID=='CDISCPILOT01', DELFN == 50) %>%
    arrange(STUDYID, USUBJID, DTTM) %>%
    select(USUBJID,DTTM,VISIT,PCTPT,PCTEST,DVIDC,DV),
    title = "Study 01 Analyst-Identified Outliers")
  ) %>%
  add_listing_to_report(
    listing_table(
    broken_pk_samples %>% 
    filter(STUDYID=='CDISCPILOT02', DELFN == 50) %>%
    arrange(STUDYID, USUBJID, DTTM) %>%
    select(USUBJID,DTTM,VISIT,PCTPT,PCTEST,DVIDC,DV),
    title = "Study 02 Analyst-Identified Outliers") 
  )

 
  
```

`get_listing_info()` allows you to see the listings added to the deletion report. This can be useful when troubleshooting or comparing disposition counts to listings.`id_var` argument adds subject counts. Default is `USUBJID`. Change it to `ID` or `SUBJID` etc. based on your dataset or ignore it by setting to `none`

```{r}
get_listing_info(del_report)
```

```{r}
get_listing_info(del_report,id_var = "none") %>%
  dplyr::slice(7) %>%
  tidyr::unnest(listing)
```

## Write Listings

file path by default will be the path specified during report initialization with `build_report`. The `...` is for extra arguments to be passed to [reporter::write_report()](https://reporter.r-sassy.org/reference/write_report.html)

```{r, eval=FALSE}

# Write Listings
write_listings_report(del_report)

```
